\input{../cs164.tex}
\usepackage{amsmath, verbatim, tikz, float, dsfont}

\oddsidemargin 0in
\evensidemargin 0in
\textwidth 6.5in
\topmargin -0.5in
\textheight 9.0in
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\?}{\stackrel{?}{=}}

\begin{document}

\solution{Nikhil Unni (cs164-es), Section : Monday 3pm}{WA4}{Fall 2015}
\pagestyle{myheadings}


\begin{enumerate}
    \item
    \begin{enumerate}
        \item
        Is the following grammar LL(1)? Is it LR(1)? Explain both of your answers.
        $$F \rightarrow mv$$
        $$F \rightarrow ma$$\\

        It is not LL(1), because the grammar is not left-factored (there will be a first/first conflict on ``m''). It is LR(1) because there can be a valid parsing table which just shifts on m, and reduces on either v or a to F on a \$ in the input.

        \item
        Is the following grammar ambiguous? Is it LR(1)? Explain both of your answers.
        $$A \rightarrow Bad$$
        $$A \rightarrow Cat$$
        $$B \rightarrow r$$
        $$C \rightarrow r$$\\

        It is not ambiguous, just because there is only 1 parse tree each for both ``rat'' and ``rad'', the only two stings in the language. It is \textbf{not} LR(1) because there is a reduce/reduce conflict in the parsing table between $B \rightarrow r$ and $C \rightarrow r$. This can be remedied by changing $A \rightarrow Cat$ to $A \rightarrow Bat$, and eliminate the nonterminal C.
    \end{enumerate}

    \item Consider the following grammar with start symbol S:
    $$S \rightarrow Dx$$
    $$D \rightarrow pwD | n$$

    \begin{enumerate}
        \item Complete the DFA skeleton.\\

        I drew it on the back.\\
        \item Is the grammar LR(1)? LR(2)? LL(1)?\\

        The grammar is LR(1) because we have no conflicts in our DFA. Because the grammar is LR(1) it \textbf{has} to be LR(2) as well, (and, generally, LR(k), $k > 0$). The grammar is also LL(1) because it has no left recursion and is completely left factored.\\
        \item Use your DFA to parse pwpwpwnx.
            \begin{center}
            \begin{tabular}{c c c}
               Stack & Input & Action\\
               \hline
                 & $\triangleright$pwpwpwnx$\$$ & shift 7\\
                 pwpwpwn & pwpwpwn$\triangleright$x$\$$ & reduce $D \rightarrow n$\\
                 pwpwpwD & pwpwpwn$\triangleright$x$\$$ & reduce $D \rightarrow pwD$\\
                 pwpwD & pwpwpwn$\triangleright$x$\$$ & reduce $D \rightarrow pwD$\\
                 pwD & pwpwpwn$\triangleright$x$\$$ & reduce $D \rightarrow pwD$\\
                 D & pwpwpwn$\triangleright$x$\$$ & shift\\
                 Dx & pwpwpwnx$\triangleright$$\$$ & reduce $S \rightarrow Dx$\\
                 S & pwpwpwnx$\triangleright$$\$$ & accept
            \end{tabular}
            \end{center}
    \end{enumerate}
\end{enumerate}
\end{document}
